The Sloan Digital Sky Survey Spectroscopic Reduction Pipeline

### 2006-04-25 doc/paper/paper2d.tex
Added crude figure of spectrograph layout
small caption change
Update M67 plots.

A plate was drilled centered on the open cluster M67 with the
specific idea of testing the SDSS velocity precision.
130 stars were targetted from the radial velocity catalog
of Mathieu \etal\ (1986), falling in the magnitude range of 8 to 14.
The 19 stars brighter than $m=11$ were not plugged, since they
would have saturated the detectors.  This was plate 321 MJD 51612.

@@

Below, I've listed the vacuum wavelength of some transitions:
\begin{verbatim}
 Air       Vacuum
 --------  --------
 4861.363  4862.721  H_beta
 4958.911  4960.295  [O III]
 5006.843  5008.239  [O III]
 6548.05   6549.86   [N II]
 6562.801  6564.614  H_alpha
 6583.45   6585.27   [N II]
 6716.44   6718.29   [S II]
 6730.82   6732.68   [S II]
\end{verbatim}
The wavelengths are shifted such that measured velocities are relative
to barycentric at the mid-point of each 15-minute exposure.

@@

\begin{figure}[tb]
\figurenum{1}
\epsscale{1.0}
\plotone{plate_stats.ps}
\figcaption{Observational summary of the 1656 publicly released
SDSS spectroscopic
plates observed between MJD 2451600 to 2453600.  Each plate is
represented by 3 symbols showing the median RMS guiding residuals
in the 11 guide fibers in arcseconds (diamonds), 
the median seeing of the 11 guide star PSFs in arcseconds (stars), 
and the effective (S/N)$^2$ of the plate observation
normalized to a 1 hour exposure.  Vertical lines show the date of July 15 across
the years of 2000 to 2005.
\label{fig1}
}
\end{figure}

\begin{figure}[tb]
\figurenum{2}
\epsscale{1.0}
\plotone{spDiag2d-0412-52258.ps_1}
\figcaption{A typical example of the common flat-field spectra, $SF$,
for the b1 camera observations of plate 412.  The solid line is the best-fit
B-spline function, and the red points are rejected 2-sigma outliers. 
\label{fig2}
}
\end{figure}

\begin{figure}[tb]
\figurenum{7}
\epsscale{0.9}
\plotone{plot_disp.ps}
\figcaption{Measurements of spectral resolution in the four SDSS 
spectrograph cameras.  For each camera, we should the median
FWHM arc-line widths as a function of wavelength as 
solid lines for cameras b1 (cyan) and r1 (orange),
dashed lines for cameras b2 (blue) and r2 (red).
The range of resolution is shown as error bars representing
the 10\% to 90\% ranges in each camera over all fibers and
all plates through DR5.
\label{fig7}
}
\end{figure}

\begin{figure}[tb]
\figurenum{8}
\epsscale{1.0}
\plotone{spDiag2d-0412-52258_156.eps}
\figcaption{An example of 1-d sky subtraction for camera r1 on SDSS
plate 412.  The top panel shows data from 16 sky fibers corrected to airmass
unity with vertical extent equal to 67\% errors.  The solid red line shows
the best fit B-spline, with only wavelength as an independent variable.
The bottom panel shows the residual of the fit in flux counts in the sky 
spectra.
\label{fig8}
}
\end{figure}

\begin{figure}[tb]
\figurenum{9}
\epsscale{1.0}
\plotone{spDiag2d-0412-52258.ps_153}
\figcaption{An example of the wavelength corrections to match expected 
sky emission line wavelength centers in a single r1 exposure.  The position
residuals in spectral pixels are shown as measurements of 12 sky 
emission lines for all 320 fibers.  The red positions show the initial 
predicted positions based on the arc-lamp wavelength solutions, and the green
positions are based on the updated solution after shifting and stretching to
match the observed sky lines positions.  The residuals can be compared to a
full pixel in the top left corner, and are much smaller than 5\%.
\label{fig9}
}
\end{figure}

>>>>

A plate was drilled centered on the open cluster M67 with the
specific idea of testing the SDSS velocity precision.
This was plate 321 observed on MJD 51612.  
130 stars were targetted from the radial velocity catalog
of Mathieu \etal\ (1986), falling in the magnitude range of 8 to 14.
The 19 stars brighter than $m=11$ were intentionally not plugged, since they
would have saturated the detectors, and another 8 stars appear not
to have been plugged.  SDSS velocities for these stars were determined
by fitting these spectra to all stars in the Elodie-2 stellar catalog,
and choosing the best-fit Elodie star using $\chi^2$ minimization.  (ref???)
The derived veocites have a median offset of $-5.5\kms$, with a sigma-clipped
dispersion of $3.0 \kms$ (see Figure \ref{fig_m67}).
There were three very discrepent velocities. (look these up???)
% Fiber 224 at RA= 132.40396 Dec= 11.484270 v(CAT)= 13.30 v(Elodie)= 100.50532
% Fiber 301 at RA= 132.01817 Dec= 12.167990 v(CAT)= 12.10 v(Elodie)= 104.35960

Our expectation is that this plate represents a worst-case scenario
for the zero-points in the wavelength scales, and therefore velocties.
The exposure times were 30, 60 and 120 sec for this plate, as opposed
to typical exposure times of 900 to 1500 sec on SDSS main survey plates.
With the shorter exposures, there is not as much signal in the sky lines
for computing the instrument flexure between the arc lamp spectra and
the science exposures.

@@

Below are listed the vacuum wavelength of some transitions:
\begin{verbatim}
 Air       Vacuum
 --------  --------
 4861.363  4862.721  H_beta
 4958.911  4960.295  [O III]
 5006.843  5008.239  [O III]
 6548.05   6549.86   [N II]
 6562.801  6564.614  H_alpha
 6583.45   6585.27   [N II]
 6716.44   6718.29   [S II]
 6730.82   6732.68   [S II]
\end{verbatim}
The wavelengths are shifted such that measured velocities are relative
to barycentric at the mid-point of each 15-minute exposure.

@@

\begin{figure}
\figurenum{1}
\epsscale{1.0}
\plotone{plate_stats.ps}
\figcaption{Observational summary of the 1656 publicly released
SDSS spectroscopic
plates observed between MJD 2451600 to 2453600.  Each plate is
represented by 3 symbols showing the median RMS guiding residuals
in the 11 guide fibers in arcseconds (diamonds), 
the median seeing of the 11 guide star PSFs in arcseconds (stars), 
and the effective (S/N)$^2$ of the plate observation
normalized to a 1 hour exposure.  Vertical lines show the date of July 15 across
the years of 2000 to 2005.
\label{fig1}
}
\end{figure}

\begin{figure}
\figurenum{2}
\epsscale{0.9}
\plotone{layout.ps}
\figcaption{Schematic layout of one of the SDSS spectrographs, from the 
SDSS Project Book (http://www.astro.princeton.edu/PBOOK/spectro/spectro.htm).
Four main components are labeled: A) the slithead, B) the collimator, 
C) the beamsplitter, D) the grisms.
\label{fig2}
}
\end{figure}

\begin{figure}[tb]
\figurenum{3}
\epsscale{1.0}
\plotone{spDiag2d-0412-52258.ps_1}
\figcaption{A typical example of the common flat-field spectra, $SF$,
for the b1 camera observations of plate 412.  The solid line is the best-fit
B-spline function, and the red points are rejected 2-sigma outliers. 
\label{fig3}
}
\end{figure}

\begin{figure}[tb]
\epsscale{0.9}
\plotone{plot_disp.ps}
\figcaption{Measurements of spectral resolution in the four SDSS 
spectrograph cameras.  Shown for each camera is the median
FWHM arc-line widths as a function of wavelength as 
solid lines for cameras b1 (cyan) and r1 (orange),
dashed lines for cameras b2 (blue) and r2 (red).
The range of resolution is shown as error bars representing
the 10\% to 90\% ranges in each camera over all fibers and
all public data release plates through DR5.
\label{fig_disp}
}
\end{figure}

\begin{figure}
\epsscale{0.9}
\plotone{plot_m67.ps}
\figcaption{
SDSS velocities of 103 stars in or near the open cluster M67
compared to published velocities in Mathieu \etal\  (1986).
There is an offset of $5.0\kms$, and a sigma-clipped dispersion
of $3.0\kms$.
\label{fig_m67}
}
\end{figure}

\begin{figure}
\figurenum{8}
\epsscale{1.0}
\plotone{spDiag2d-0412-52258_156.eps}
\figcaption{An example of 1-d sky subtraction for camera r1 on SDSS
plate 412.  The top panel shows data from 16 sky fibers corrected to airmass
unity with vertical extent equal to 67\% errors.  The solid red line shows
the best fit B-spline, with only wavelength as an independent variable.
The bottom panel shows the residual of the fit in flux counts in the sky 
spectra.
\label{fig8}
}
\end{figure}

\begin{figure}
\figurenum{9}
\epsscale{1.0}
\plotone{spDiag2d-0412-52258.ps_153}
\figcaption{An example of the wavelength corrections to match expected 
sky emission line wavelength centers in a single r1 exposure.  The position
residuals in spectral pixels are shown as measurements of 12 sky 
emission lines for all 320 fibers.  The red positions show the initial 
predicted positions based on the arc-lamp wavelength solutions, and the green
positions are based on the updated solution after shifting and stretching to
match the observed sky lines positions.  The residuals can be compared to a
full pixel in the top left corner, and are much smaller than 5\%.
\label{fig9}
}
\end{figure}
